\documentclass{beamer}
\usepackage[utf8]{inputenc}

\usetheme{Berkeley}
\usecolortheme{default}

\setbeamertemplate{footline}[frame number]
\beamertemplatenavigationsymbolsempty

\title{Course Evaluation}
\subtitle{Your feedback helps improve this course}
\author[]{IB 514: Scientific Writing}
\date[]{}

\begin{document}

% -------------------------------------------------
\begin{frame}
  \titlepage
\end{frame}

% -------------------------------------------------

\section{Content Coverage}

\begin{frame}{Content Coverage}
  \begin{itemize}
    \item Which class topics did you find most and least useful or effective?
    \bigskip
    \item Are there topics you wish had been covered or covered in greater depth?
    \bigskip
    \item Should there be more discussion of AI tools (effective uses, ethics, etc.)?
  \end{itemize}
\end{frame}

% -------------------------------------------------

\section{Format and Structure}

\begin{frame}{Format and Structure}
  \begin{itemize}
    \item Which class activities did you find most and least useful or effective?
    \bigskip
    \item Should the course include more activities, more writing time, more lecture/discussion classes, or is the current balance about right?
    \bigskip
    \item How might I motivate more discussion in class?\\
      \smallskip
      {\small (How did you use/engage with the slides? Did you review them before class?)}
  \end{itemize}
\end{frame}

% -------------------------------------------------

\section{Specific Course Features}

\begin{frame}{Specific Course Features}
  \begin{itemize}
    \item Was pairing up with an accountabilibuddy helpful for your productivity and motivation?
      Should this be continued in future offerings?
  \end{itemize}
\end{frame}

% -------------------------------------------------

\section{Feedback \& Accountability}

\begin{frame}{Feedback \& Accountability}
  \begin{itemize}
    \item Did you receive enough individual feedback on your writing --- from the instructor, peers, or both --- to make meaningful revisions?
    \bigskip
    \item Would additional accountability requirements (e.g., graded exercises, mandatory manuscript benchmarks) have been a useful motivator for greater productivity?
  \end{itemize}
\end{frame}

% -------------------------------------------------

\end{document}
